An Empirical Study of Invasion and Resistance for Iterated Prisoner's Dilemma Strategies

### 2017-05-15 main.tex
Make minor adjustments.

As well as noting that the memory length and complexity of these strategies are
much greater than one, it is interesting to note that none of them are akin to
memory one strategies. Only one is stochastic although close inspection of the
source code of PSO Gambler shows that it makes stochastic decisions rarely, and
is functionally very similar to its deterministic cousin EvolvedLookerUp. Apart
from TF3 in \(N=3\), the finite state machines trained specifically for Moran
processes do not appear in the top 5: whereas strategies trained for tournaments do.
This is due to the nature of invasion: most of the opponents will initially be different
strategies. The next section will consider the converse situation.

@@

The strategies trained to maximize payoff in head-to-head matches are generally
cooperative and are effective invaders.
Combined with the fact that handshaking strategies are stronger resisters,
this suggests that while maximizing individual payoff can lead to the evolution
of cooperation, these strategies are not the most evolutionarily stable
in the long run. A strategy with a handshaking mechanism is still capable of
invading and is more resistant to subsequent invasions. Moreover, the
best resistor of the payoff maximally trained strategies (EvolvedLookerUp1_1_1),
which always defects if the opponent defects in the first round, is effectively
employing a one-shot handshake of C. These insights also suggest that a strategy
aware of the population distribution could choose to become a handshaker at
a critical threshold and use a different strategy for invasion when in the
minority. However this information was not available to our strategies.

We did not attempt other objective functions that may serve to select for both
invasion and resistance better than training at a starting population of
$(N/2, N/2)$. Nevertheless our results suggest that there is not much room for
improvement. Any handshake more sophisticated than C necessarily involves
a defection. (A handshake consisting of a longer sequence of cooperations is
effectively a grudger.) For TF3 or EvolvedLookerUp1_1_1 to become better resistors
they need a longer or more strict handshake. But if this handshake involves
a defection then likely the invasion ability is diminished for $N > 2$: the top
invaders for larger $N$ are nice strategies that do not defect before their
opponents. This is because good invaders still need to cooperate with themselves,
and so in the absense of a handshake mechanism or knowledge of the population
distribution, a strategy must be generally cooperative. Aggressive strategies
are only effective invaders for the smallest $N$, dropping down dramatically
as the population size increases.

>>>>

As well as noting that the memory length and complexity of these strategies are
much greater than one, it is interesting to note that none of them are akin to
memory one strategies. Only one is stochastic although close inspection of the
source code of PSO Gambler shows that it makes stochastic decisions rarely, and
is functionally very similar to its deterministic cousin Evolved Looker Up. Apart
from TF3 in \(N=3\), the finite state machines trained specifically for Moran
processes do not appear in the top 5: whereas strategies trained for tournaments do.
This is due to the nature of invasion: most of the opponents will initially be different
strategies. The next section will consider the converse situation.

@@

The strategies trained to maximize payoff in head-to-head matches are generally
cooperative and are effective invaders.
Combined with the fact that handshaking strategies are stronger resisters,
this suggests that while maximizing individual payoff can lead to the evolution
of cooperation, these strategies are not the most evolutionarily stable
in the long run. A strategy with a handshaking mechanism is still capable of
invading and is more resistant to subsequent invasions. Moreover, the
best resistor of the payoff maximally trained strategies (Evolved Looker Up
1\_1\_1),
which always defects if the opponent defects in the first round, is effectively
employing a one-shot handshake of C. These insights also suggest that a strategy
aware of the population distribution could choose to become a handshaker at
a critical threshold and use a different strategy for invasion when in the
minority. However this information was not available to our strategies.

We did not attempt other objective functions that may serve to select for both
invasion and resistance better than training at a starting population of
$(N/2, N/2)$. Nevertheless our results suggest that there is not much room for
improvement. Any handshake more sophisticated than C necessarily involves
a defection. (A handshake consisting of a longer sequence of cooperations is
effectively a grudger.) For TF3 or EvolvedLookerUp1\_1\_1 to become better resistors
they need a longer or more strict handshake. But if this handshake involves
a defection then likely the invasion ability is diminished for $N > 2$: the top
invaders for larger $N$ are nice strategies that do not defect before their
opponents. This is because good invaders still need to cooperate with themselves,
and so in the absense of a handshake mechanism or knowledge of the population
distribution, a strategy must be generally cooperative. Aggressive strategies
are only effective invaders for the smallest $N$, dropping down dramatically
as the population size increases.
